## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}

```

We can also add else if statements in the middle if we need more options.

```{r}

```

Neat! We can also nest conditionals just like we do in Python.

```{r}

    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}

```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}

```

We can loop over more than just lists:

```{r}

```

Like Python, we have while loops in R:

```{r}

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}

```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}

```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}

```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}

```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
    
    #1
  divisors <- function(n) {
  if (is.numeric(n) && n > 0 && length(n) == 1) {
    divisors <- 0
    for (i in 1:sqrt(n)) {
      if (n %% i == 0) {
        # If i is a divisor, both i and n/i are divisors
        divisors <- divisors + ifelse(i == n/i, 1, 2)
      }
    }
    return(divisors)
  } else {
    stop("Input must be a single positive numeric value.")
  }
  }
    
    
    #2
    bouncingBall <- function(h, bounce, window) {
      if (is.numeric(h) && is.numeric(bounce) && is.numeric(window) &&
          h > 0 && bounce > 0 && bounce < 1 && window < h) {

        count <- 0
        while (h > window) {
          count <- count + 1
          h <- h * bounce

          # Check if the ball can be seen after bouncing
          if (h > window) {
            count <- count + 1
          }
        }

        return(count)
      } else {
        return(-1)
      }
    }
    
    
    # 3
    
    your_order <- function(inputString){
   if (nchar(inputString) == 0) {
      return("")
    }

    words <- strsplit(inputString, " ")[[1]]
    numbers <- as.numeric(gsub("\\D", "", words))
    sorted_indices <- order(numbers)

    sorted_words <- words[sorted_indices]
    sorted_string <- paste(sorted_words, collapse = " ")

    return(sorted_string)
    }
    
    
    #4
    
    revRot <- function(str, sz) {
    if (sz <= 0 || nchar(str) == 0 || sz > nchar(str)) {
        return("")
      }
    
      chunks <- substring(str, seq(1, nchar(str), by = sz), seq(sz, nchar(str), by = sz))
      modified_chunks <- sapply(chunks, function(chunk) {
        if (sum(as.numeric(strsplit(chunk, "")[[1]])^3) %% 2 == 0) {
          return(paste(rev(strsplit(chunk, "")[[1]]), collapse = ""))
        } else {
          rotated_chunk <- substring(chunk, 2, nchar(chunk))
          rotated_chunk <- paste(rotated_chunk, substr(chunk, 1, 1), sep = "")
          return(rotated_chunk)
        }
      })
    
      return(paste(modified_chunks, collapse = ""))
    }
    
    
    #5
    angle <- function(n) {
       if (n > 2) {
          total_sum <- (n - 2) * 180
          return(total_sum)
        } else {
          stop("Number of sides must be greater than 2.")
        }
      }

    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
fizz_buzz <- function(number) {
  if (is.numeric(number) && length(number) == 1) {
    if (number %% 3 == 0 && number %% 5 == 0) {
      return("FizzBuzz")
    } else if (number %% 3 == 0) {
      return("Fizz")
    } else if (number %% 5 == 0) {
      return("Buzz")
    } else {
      return(as.character(number))
    }
  } else {
    stop("Input must be a single numeric value.")
  }
}

# Example usage:
result <- fizz_buzz(15)
print(result)  # Output: "FizzBuzz"

    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.
marioStairs <- function(length, height) {
  if (length > 0 && height > 0) {
    for (i in 1:height) {
      cat(strrep("X", i), "\n")
    }
  } else {
    stop("Length and height must be positive integers.")
  }
}

# Example usage:
marioStairs(4, 4)

    ```
